---
title: "IDA Homework - Week 1"
author: "Adam"
output: pdf_document
---

# Task 1: Reproducing the table

## The data set I: creating a table in RMarkdown

### Overview of variables in the data set

**Variable** | **Variable type** | **Values**
:-----------:|:-----------------:|:----------------------------
brain regions | nominal | (Rostral Middle Frontal, Precentral, Lateral Occipital, Transverse Temporal, Temporal Pole)
surface area | volume | (5941.8, 4718.8, 4672.9, 799.48, 443.3)
thickness | metric | (2.59, 2.74, 2.3, 2.52, 3.66)
hemisphere | ordinal | (L, L, L, R, R)

# Task 2: Loading packages

## Loading relevant packages
```{r, warning=FALSE, message=FALSE, include=FALSE}
library(tidyverse)
```

# Task 3: Creating a tibble
## The data set II: Creating a tibble in R

```{r, include=FALSE}
brain_data <- tibble(
  brain_regions = factor(c("Rostral Middle Frontal", "Precentral", "Lateral Occipital", "Transverse Temporal", "Temporal Pole")),
  surface_area = c(5941.8, 4718.8, 4672.9, 799.48, 443.3),
  thickness = c(2.59, 2.74, 2.3, 2.52, 3.66),
  hemisphere = c("L", "L", "L", "R", "R")
)
brain_data
```

# Task 4: Altering data

Changing "Transverse temporal" to "Banks superior temporal"

```{r, include=FALSE}
brain_data$brain_regions <- fct_recode(brain_data$brain_regions, "Banks Superior Temporal" = "Transverse Temporal") %>%
  print()
```

# Task 5: Indexing

```{r, echo=FALSE}
# number method
brain_data[2, 3] %>%
  print()

# index by name
brain_data[brain_data$brain_regions == 'Precentral', "thickness"] %>%
  print()
```

# Task 6: Considering the data set
## The data set III: Creating a tibble in R
```{r}
brain_data2 <- tibble(
  brain_regions = factor(c("Rostral Middle Frontal", "Precentral", "Lateral Occipital", "Transverse Temporal", "Temporal Pole")),
  surface_area = c(5941.8, 4718.8, 4672.9, 799.48, 443.3),
  thickness = c(2.59, 2.74, 2.3, 2.52, 3.66),
  hemisphere = c("L", "L", "L", "R", "R"),
  volume = surface_area * thickness
) %>% print()
```

# Task 7: Why wouldn't this code work for a dataframe
Because dataframes don't allow for flexible creation (e.g. dynamic multiplying elements).

# Task 8: Store specific code
```{r}
regions_v1 <- brain_data2$brain_regions %>% print()
```

# Task 9: Datatype
```{r}
typeof(regions_v1)
```

# Task 10: How many variables
```{r}
length(regions_v1)
```

# Task 11: How many elements total in hemisphere
```{r}
brain_data2$hemisphere %>% unique() %>% length() %>% print()
```

# Task 12: Describing code
Code is meant to find what brain regions of the data set are in the right hemisphere.

# Task 13: Using a function for volume
```{r}
# extract surface_area from dataset
x <- brain_data2$surface_area
# extract thickiness from dataset
y <- brain_data2$thickness
# create a function that multiplies them together 
volume_calc <- function(area, thick) {
  return(area*thick)
}
# call the function
volume_calc(x, y)

# create comparative list
volume_compare <- list(
  volume1 <- brain_data2$volume,
  volume2 <- volume_calc(x, y)
) %>% print()
```

# Task 14: Extracting 
```{r}
# listing true values
volume_true <- c(17439, 14351, 12150, 23886, 2280.1)

# extract calculated from list
volume_calculated <- volume_compare$volume1 %>% print()
# for some reason this is borked and I cannot for the life of me work out why

# function that takes two variables and returns the difference
volume_diff <- function(vol1, vol2){return(vol1-vol2)}
```

# Task 15: Tibbles and data
```{r}
brain_data3 <- tibble(
  brain_regions = regions_v1,
  calculated_volume_diff = volume_diff(volume_true,volume_calculated)
) %>% print()
```
